### 2014-09-19 text/Chapter04.tex
Finish conclusions and tweaks (Chapter 5)

For top-pointed coeffect algebra (using Lemma~\ref{thm:cbn-substitution-top}), the proof is similar
to the one in Theorem~\ref{thm:cbv-reduction}, using the facts that $\cclrd{s} \cleq \cunit$
and $r \,\czip\, t \;=\; r \,\cpar\, t$.
For bottom-pointed coeffect algebra, consider the typing derivation for the term
$(\lambda x.e)\,v$ before reduction:
%
\begin{equation*}
\inference
  { \inference
      {\coctx{\Gamma, x:\tau_s}{\cclrd{r}} \vdash e_r : \tau_r}
      {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e_r : \tau_s \xrightarrow{\cclrd{r}} \tau_r } &
   \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s }
  {\coctx{\Gamma}{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})} \vdash (\lambda x.e_r)\,e_s : \tau_r }
\end{equation*}
%
The derivation uses the idempotence of $\czip$ in the first step, followed by the 
(\emph{app}) rule. The type of the term after substitution, using Lemma~\ref{thm:cbn-substitution-bot} is:
%
\begin{equation*}
\inference
  { \coctx{\Gamma, x:\tau_s}{\cclrd{r}} \vdash e_r : \tau_r & 
    \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s }
  { \coctx{\Gamma, x:\tau_r}{\cclrd{r}\;\cseq\;\cclrd{s}} \vdash \subst{e_r}{x}{e_s} : \tau_s }
\end{equation*}
%
From the assumptions of Lemma~\ref{thm:cbn-substitution-bot}, we know that $\cseq\,=\,\cpar$
and the operation is idempotent, so trivially:
$\cclrd{r}\;\cseq\;\cclrd{s} = \cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})$
\end{proof}

@@

\begin{theorem}[$\eta$-expansion]
\label{thm:cbn-substitution-bot}
In a bottom-pointed flat coeffect calculus with an algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$ 
where $\czip = \cpar$, if $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2$ 
and $e \rightarrow_\eta e'$ then $\coctx{\Gamma}{\cclrd{r}} \vdash e' : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 $.
\end{theorem}
\begin{proof}
The following derivation shows that $\lambda x.e~x$ has the same type as $e$:
\begin{equation*}
\inference
  { \inference
    { \coctx{\Gamma}{\cclrd{r}} \vdash e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 &
      \coctx{x:\tau_1}{\cunit} \vdash x:\tau_1 }
    { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\cpar\; (\cunit \,\cseq\, \cclrd{s}) } \vdash e~x : \tau_2 } }
  { \inference
    { \inference
      { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\cpar\; \cclrd{s} } \vdash e~x : \tau_2 }
      { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\czip\; \cclrd{s} } \vdash e~x : \tau_2 } }
    { \coctx{\Gamma}{ \cclrd{r} } \vdash \lambda x.e~x : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 } }
\end{equation*}
\end{proof}

>>>>

For top-pointed coeffect algebra (using Lemma~\ref{thm:cbn-substitution-top}), the proof is similar
to the one in Theorem~\ref{thm:cbv-reduction}, using the facts that $\cclrd{s} \cleq \cunit$
and $r \,\czip\, t \;=\; r \,\cpar\, t$.
For bottom-pointed coeffect algebra, consider the typing derivation for the term
$(\lambda x.e_r)\,e_s$ before reduction:
%
\begin{equation*}
\inference
  { \inference
      {\coctx{\Gamma, x:\tau_s}{\cclrd{r}} \vdash e_r : \tau_r}
      {\coctx{\Gamma}{\cclrd{r}} \vdash \lambda x.e_r : \tau_s \xrightarrow{\cclrd{r}} \tau_r } &
   \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s }
  {\coctx{\Gamma}{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})} \vdash (\lambda x.e_r)\,e_s : \tau_r }
\end{equation*}
%
The derivation uses the idempotence of $\czip$ in the first step, followed by the 
(\emph{app}) rule. The type of the term after substitution, using Lemma~\ref{thm:cbn-substitution-bot} is:
%
\begin{equation*}
\inference
  { \coctx{\Gamma, x:\tau_s}{\cclrd{r}} \vdash e_r : \tau_r & 
    \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s }
  { \coctx{\Gamma, x:\tau_r}{\cclrd{r}\;\cseq\;\cclrd{s}} \vdash \subst{e_r}{x}{e_s} : \tau_s }
\end{equation*}
%
From the assumptions of Lemma~\ref{thm:cbn-substitution-bot}, we know that $\cseq\,=\,\cpar$
and the operation is idempotent, so trivially:
$\cclrd{r}\;\cseq\;\cclrd{s} = \cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})$
\end{proof}

@@

\begin{theorem}[$\eta$-expansion]
In a bottom-pointed flat coeffect calculus with an algebra $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$ 
where $\czip = \cpar$, if $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2$ 
and $e \rightarrow_\eta e'$ then $\coctx{\Gamma}{\cclrd{r}} \vdash e' : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 $.
\end{theorem}
\begin{proof}
The following derivation shows that $\lambda x.f~x$ has the same type as $f$:
\begin{equation*}
\inference
  { \inference
    { \coctx{\Gamma}{\cclrd{r}} \vdash f : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 &
      \coctx{x:\tau_1}{\cunit} \vdash x:\tau_1 }
    { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\cpar\; (\cunit \,\cseq\, \cclrd{s}) } \vdash f~x : \tau_2 } }
  { \inference
    { \inference
      { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\cpar\; \cclrd{s} } \vdash f~x : \tau_2 }
      { \coctx{\Gamma, x:\tau_1}{\cclrd{r} \;\czip\; \cclrd{s} } \vdash f~x : \tau_2 } }
    { \coctx{\Gamma}{ \cclrd{r} } \vdash \lambda x.f~x : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 } }
\end{equation*}
\end{proof}
